---
title: "Counting Calories"
author: "Homer White"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
library(tidyverse)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  error = TRUE,
  tidy = FALSE,
  size = "small"
)
```


## A Quick Solution

Get the data:

```{r}
input <- readLines("input.txt")
```

Make it numeric:

```{r}
calories <- parse_number(input)
```

```{r}
df <-
  data.frame(
    calories = calories
  )
```

Every row  (except for the rows with `NA`) stands for an item in the rucksack of an elf.

We would like a new column that has zeroes where there are calories, and 1s where there are `NA`-values.

```{r}
df <-
  df %>% 
  mutate(is_dividing_line = ifelse(
    is.na(calories),
    1, 0
    )
  )

head(df, n = 5)
```

We would like to be able to group the items together into groups corresponding to each elf.  For this we use `cumsum()`.  Run `?cumsum` before proceeding.

```{r}
df <-
  df %>% 
  mutate(elf = cumsum(is_dividing_line))
head(df, n = 20)
```

The elf ids are 0, 1, 2, and so on.


We do not need the `NA`-values any more, so let's drop all the rows where `calories` is `NA`:

```{r}
df <- 
  df %>% 
  drop_na()
```

Now we have a tidy data table, and we are ready to wrangle in the usual **dplyr** way:

```{r}
rucksack_sums <-
  df %>% 
  group_by(elf) %>% 
  summarize(total_calories = sum(calories))
head(rucksack_sums)
```

We just want the biggest sum:

```{r}
rucksack_sums %>% 
  slice_max(order_by = total_calories, n = 1) %>% 
  pull(total_calories)
```


Of course you can pipe it all, like this:

```{r}
df <-data.frame(
    calories = readLines(
      con = "input.txt"
    ) %>% 
      parse_number()
  ) %>% 
  mutate(is_dividing_line = ifelse(
    is.na(calories),
    1, 0
  )) %>% 
  mutate(elf = cumsum(is_dividing_line)) %>% 
  drop_na() %>% 
  group_by(elf) %>% 
  summarize(total_calories = sum(calories))
```

Then for Part One:

```{r}
df %>% 
  slice_max(n = 1, order_by = total_calories) %>% 
  pull(total_calories)
```

For Part Two:

```{r}
df %>% 
  slice_max(n = 3, order_by = total_calories) %>% 
  pull(total_calories) %>% 
  sum()
```


## A Loopy Approach

Begin as before:

```{r}
input <- readLines("input.txt")
calories <- parse_number(input)
```

Now the loop:

```{r}
rucksack_sums <- numeric()
current_item <- 1

while(current_item <= length(calories)) {
  pack_sum <- 0
  while(!(is.na(calories[current_item]))) {
    pack_sum <- pack_sum + calories[current_item]
    current_item <- current_item + 1
  }
  rucksack_sums <- c(rucksack_sums, pack_sum)
  ## move past the NA to the next elf:
  current_item <- current_item + 1
}
```

Then sort the sums:

```{r}
sorted_sums <-
  sort(rucksack_sums, decreasing = TRUE)
```

Answer to Part One:

```{r}
sorted_sums[1]
```

Answer to Part Two:

```{r}
sorted_sums[1:3] %>% sum()
```








